To Airbnb?: A look at Long vs Shorm Term Leasing

### 2018-11-06 papers/full_paper.Rmd
Revision 2, edits.

The growth of Airbnb and other short-term rental platforms have presented absentee owners of residential properties with a choice of leasing strategy: traditional long-term rental or a short-term approach ("Airbnb-ing").  In this paper we identify those situations -- location, structure type and property characteristics -- that lead to the highest likelihood of favoring a short-term strategy over a long term one. We also find that hosting policies matter, suggesting that even the right property may need the right owner(s) or strategy to make short-term rental the more profitable approach.

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? Are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one? Do host policies matter?

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to the urbanized areas of Melbourne. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachement to the property.  While we analyze individual properties, we acknowledge that there are many unobserved condiseration and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

Our results show that the feasibility of a greater revenues through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor should opt for a short-term leasing strategy is ''it depends''.

@@

Much of the existing work focuses on price or rent formation in second home markets.  Specifically, the impacts of amenities such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water], proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  Host policies can also matter as @benjamin2001value find that smoking prohibitions influence the weekly rental rates for vacation homes.

@@

Though issues surrounding second homes have been primarily treated as a rural phenomenon for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

We approach answering the question of whether it is more profitable to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a residential investment property.  More specifically, we examine the following case:

* An investor purchases a property in Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* They do not intend to inhabit the home themselves. 
* The choices are 1) traditional year long rental (long-term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are ignored as these concerns are influenced by owner situations, not property characteristics or hosting policies.  
* As most long-term rentals in Melbourne use a leasing agency, we will assume use of an agency.

Australia implements a particular tax policy, often termed `negative gearing', that allows for an income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in our analysis here.

@@

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we attempted to also estimate short-term rental revenue for long-term rental properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform. On Airbnb, in addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the Airbnb data (paid separately and in addition to the nightly rate) they can also be ignored in the comparative analysis below.

@@

Data for this study were gathered from two different sources, one for the short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites.^[In fact, our choice of study area was done so to specifically minimize the number of potential traditional vacation homes in our sample.] The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property-level and nightly-per-property. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period.^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  The nightly data indicate the advertised price, the occupancy status and a unique reservation identification.

@@

As long-term rentals do not offer Private Room or Shared Room options^[Our long-term rental dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term revenues, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter does remove about 46% of the short-term data, however, given the research question it is unavoidable.

@@

Both anecdotal evidence and Airbnb's own marketing campaigns suggest that there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).  For the purposes of this research we are focused on Profit Seekers; those owner who would be faced with deciding between a long- and short-term leasing strategy.

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into three categories based on the original two variables considered: block percentage and number of blocked periods. Profit Seekers are those host who have blocked percentages less than 25\% of the time.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users we consider as Unknown as they may be multiple plaform users (who are profit seeking), very opportunistic sharers or driven by some other, unknown motivation.  Profit Seekers make up 43\% of the hosts, with Opportunistic Sharers capturing 19\% and the remainder 'Unknown' category contains just over 38\% of all properties.  To best approximate the scenario considered in our research question, we keep only the Profit Seekers in our dataset. See Figure 2 for a breakdown of these group by the two variables.

@@

Using the two remaining broad submarkets (Inner and Middle) we have created a four sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations.  The boundary of the suburbs meeting these criteria compose our study area (Figure 3). The location of the short- and long-term rentals, colored by submarket, are shown in Figures 4 and 5.

@@

![Short Term Observations](C:/Code/research/airbnbMelbourne/images/str_loc_map.png)

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map (Figure 2), the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban Airbnb market is somewhat thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

Second, many of the short-term properties have a significant number of blocked days (up to 255), days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 20\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case, the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

2. Potential ($REV_p$)}: Observed bookings extrapolated to a full year plus imputed bookings during observed blocked days.  Our attempt here is to represent the potential revenue if the property had been listed all year long.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance and then applying that relative ratio to the overall market performance on the extrapolated and blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Potential' revenue metric to be a better estimate of market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 7). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Actual short-term revenues range from neglible to highs of greater than \$60,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is shifted slightly to the right due to our time extrapolation and blocked days imputation.

While this simple analysis does not control for location and product differences, it does highlight the fact that some Airbnb owners are likely making greater revenues than in long-term rentals, but that most are likely not. The high number of very low yield Airbnb properties suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

To limit the potential impacts of systematic bias from omitted variables or misspecification in our hedonic model on later analyses, we've also used a matched-pairs or clustering approach as the other half of our ensemble.  In this method, we first group properties based on structure type, bedroom and bathroom combination and then find the geographically nearest 7 long-term observations. For example, all 2-bed, 2-bath apartments are grouped.  For each short-term rental in this group we find the nearest 7 comparables (geographically), many of which are in the same building.  We compute the median of these seven comparables as our matched pairs estimation.

We then average the hedonic model and matched pairs prediction for each short term property to create an estimate of its long-term rental rate.

@@

Overall, 57.3%^[If we just look at actual revenue, ignoring blocked days and properties with less than a full 12 months of listing, the figure is 42.5%] of the 2,415 short-term rental properties in our sample were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

We specify three, increasingly complex, logistic regression models. We use the likely revenue (not potential) as the dependent variable here, saving an anaylsis of potential revenue for the subsequent sensitivity tests.

@@

For Model 2 we add the five geographic submarkets, $SubMrkt$ to better control for geographic variation in the probabilities.

@@

where $GPR$ is the total number of guests allowed per bedroom^[Hosts set the total number of guests.  We have normalized this by bedroom to avoid co-linearity in our model with the bedroom variables.], $MinStay$ is the minimum number of nights that a stay may have and $CancelPolicy$ is the difficulty with which a booking may be cancelled.  We have grouped the cancellation policies into Flexible (24 hour notice), and Strict (those requiring 5 or more days notice, often with less than 100% refunds).

@@

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 2 bath combinations. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit.

@@

Finally, in Model 3 we add in hostpolicies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference. Strict cancellation may be a luxury available only for the best and most highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform a sensitivity test. The most influential assumptions that we make are in regards to; 1) unobserved costs of short term leasing (furnishings, additional maintenance, turnover, etc.) and that the properties in the two markets in our datasets are equivalent in terms of unobserved variables, such as quality and condition.  To approximate how these assumption may be influencing our results, we re-estimating our models under two new conditions:

@@

Table 3 shows the results of our sensitivity testing, with the original model results shown for comparison.  For the five variables with the strongest relationship in the original model -- House, City, Suburban, Guests Per Bedroom and Flexible Cancel -- the signs of the coefficient remain identical across the sensitivity tests and the variation in magnitudes are small, relatively speaking.   A few of the originally less significant variables do show relatively larger deviations in coefficient magnitude, but in general, the models appear to be stable across the sentivity tests.  Do note that the intercept terms does change across models, which would be expected as this represents the base probability of short-term preference which we'd expect to move in concert with our changing assumptions

@@

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, similar to those from Profit Seeking short term rentals, however, the long-term revenues are much less variable (Figure 7).  The long right tail to the distribution of short-term revenues suggesting that some owners are doing better, and occasionally much better, in the short-term market than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figures 4 and 5).  Short-term leasing is much more common in the central areas of the city, with small clusters showing up in suburban beach locations.

In comparing the short-term observations with estimated long-term rents we found that just over half of our selected Airbnb properties brought in more revenue with its short-term leasing strategy than it would have with a long-term approach. As expected, this probability is not equal across space or property type.  Apartments are much more likely than houses to be more profitable as a short-term rental.  Core city properties are more likely than other properties to prefer a short-term strategy. For instance, 65\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long-term market while only 23\% of suburban houses did.

The logistic regression models (Tables 2 and 3) that we estimate show that most of the findings from the univariate analyses done earlier are validated.  Small apartments in the central city are the best bet for generating greater revenues through short-term leasing.  We also test for the impact of host behavior on the probability of short-term preference.  We find that more flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short-term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

Testing for the sensitivity of the initial results through changes two of our core assumptions -- short term leasing costs and sample similarity between LTR and STR properties -- we find the qualitative finding to be generally robust.  While the base probability of short-term preference change with changing assumptions (as expected), the sign and magnitudes of the co-variates remains relatively robust to changes in key assumption.

>>>>

The growth of Airbnb and other short-term rental platforms have presented absentee owners of urban residential properties with a choice of leasing strategy: traditional long-term rental or a short-term approach ("Airbnb-ing").  In this paper we identify those situations -- location, structure type and property characteristics -- that lead to the highest likelihood of favoring a short-term strategy over a long term one. We also find that hosting policies matter, suggesting that even the right property may need the right owner(s) or strategy to make short-term rental the more profitable approach.

@@

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating urban residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? Are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one? And, finally, do host policies matter?

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term urban residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data. To avoid confounding market dynamics caused by traditional vacation rentals (beach, ski and other recreation) we limit our study area to the urbanized areas of Melbourne. Additionally, for any individual property the choice between short and long term leasing may be influenced by individual financial situations (tax and mortgage financing) as well as non-financial concerns such as time availability and personal attachement to the property.  While we analyze individual properties, we acknowledge that there are many unobserved considerations and, therefore, we generalize our findings to market segments, locations and host policy decisions.

@@

Our results show that the feasibility of a greater revenues through a short-term rental -- what we have termed ``Short-Term Preference'' -- is dependent on property type, product, location and host policies.  In short, the answer to whether or not a owner or investor of an urban residential property should opt for a short-term leasing strategy is ''it depends''.

@@

Much of the existing work focuses on price or rent formation in amenity-driven second home markets.  Specifically, the impacts of amenities such as water access [@nelson2010valuing], water quality [@gibbs2002hedonic; @clapper2014water], proximity to ski facilities  [@soguel2008impact; @nelson2010valuing] and theme parks [@tsai2015impact] are shown to positively impact sales prices in tourism-dominated areas.  Host policies can also matter as @benjamin2001value find that smoking prohibitions influence the weekly rental rates for vacation homes.

@@

Though issues surrounding second homes have been primarily treated as a rural phenomenon for decades  [@gallent2007second], a wave of ``new urban tourism'' [@fuller2014stop] has focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many urban residential properties are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

@@

We approach answering the question of whether it is more profitable to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of an urban residential investment property.  More specifically, we examine the following case:

* An investor purchases a property in the urbanized area of Melbourne, Australia on September 1, 2015.
* They are looking to maximize the net revenue over the 12-month period from September 1, 2015 to August 31, 2016.
* They do not intend to inhabit the home themselves. 
* The choices are 1) traditional year-long rental (long-term), or: 2) short-term, nightly Airbnb-type rental.
* Only direct monetary revenues and costs are considered, others such as liquidity and inconvenience are ignored as these concerns are influenced by owner situations, not property characteristics or hosting policies.  
* As most long-term rentals in Melbourne use a leasing agency, we assume use of an agency.

Australia implements a particular tax policy, often termed `negative gearing', that allows for an income tax deduction to real estate investors who experience greater costs than revenues on investment properties.^[See @fane2004negative for a complete explanation of negative gearing.] To date, this policy only applies to long-term rental properties.  While this policy can influence the net revenue (post tax) of a property, the negative gearing policy requires knowledge of the investor's personal income situation.  As we do not have access to this information in the marketplace, we will not be considering any effects of this policy in this analysis.

@@

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we also attempted to estimate short-term rental revenue for long-term rental properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. The Airbnb platform serves as the source of our short-term rentals data. On Airbnb, in addition to the nightly rate that an occupant pays to the owner, there is also a separate 'cleaning fee' that is attached, as is a booking fee payable to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the Airbnb data (paid separately and in addition to the nightly rate) they can also be ignored in the comparative analysis below.

@@

Data for this study were gathered from two different sources, one for the short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites.^[In fact, our choice of study area was done so to specifically minimize the number of potential traditional vacation homes in our sample.] The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property-level and nightly-per-property. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period.^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  The nightly data indicate the advertised price, the occupancy status and a unique reservation identifier.

@@

As long-term rentals do not offer Private Room or Shared Room options^[Our long-term rental dataset does not include rooming houses, purpose built student accommodations and other types of rentals which may offer private or shared room options] and our purpose here is a comparison of long- and short-term revenues, we remove all short-term properties that do not lease the entire home or apartment. Unfortunately, this filter removes about 46% of the short-term data, however, given the research question it is unavoidable.

@@

Both anecdotal evidence and Airbnb's own marketing campaigns suggest that there are at least two types of hosts in the market: 1) Those that lease out their properties all the time in the name of profit maximization (Profit Seekers); and 2) Those that lease out their property very infrequently, for example when they are away on vacation (Opportunistic Sharers).  For the purposes of this research we are focused on Profit Seekers; those owners who would be faced with deciding between a long- and short-term leasing strategy.

@@

Attempts to divide the hosts up with unsupervised^[As there are no official host types we could not run more reliable supervised models.] clustering algorithms -- including the use of more variables -- failed to produce robust categories of hosts. As a more subjective, but implementable, approach we have divided the hosts up into three categories based on the block rate alone. Profit Seekers are those hosts who have blocked percentages of less than 25\%.  Opportunistic Sharers have blocked their properties at least 75\% of the time.  Within those two extremes, we consider users as "Unknown" as they may be multiple plaform users (who are profit seeking), very opportunistic sharers or driven by some other, unknown motivation.  Profit Seekers make up 41\% of the hosts, with Opportunistic Sharers capturing 23\% and the remainder 'Unknown' category contains just over 36\% of all properties.  To best approximate the scenario considered in our research question, we keep only the Profit Seekers in our dataset. See Figure 2 for a breakdown of these group by the two variables.

@@

Using the two remaining broad submarkets (Inner and Middle) we have created a four sub-market system as a starting point for spatially analyzing short and long-term urban rentals in Melbourne:

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations.  The boundary of the suburbs meeting these criteria compose our study area (Figure 3). The location of the short- and long-term rentals, colored by submarket, are shown in Figure 4.  It should be noted that the beach areas in our study are densely developed and represently, generally speaking, higher end residential suburbs, not vacation-based communities (as one might find much farther down the coast).

@@

<!-- ![Short Term Observations](C:/Code/research/airbnbMelbourne/images/str_loc_map.png) -->

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 5). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map (Figure 2), the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Suburban apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban Airbnb market is somewhat thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

Second, many of the short-term properties have a significant number of blocked days (up to 255), days in which the property was not offered on Airbnb.  Here too, there is a problem with comparing observed revenues from a property with a 20\% block rate with a hypothetical year long revenue stream, especially if the property was leased out on other portals during those blocked days.  And, even if we knew that the blocked days were simply due to the host wanting to use the property themselves during that time, this deduction in revenue should not necessarily apply to the hypothetical investor in our examined scenario.  In short, in either case the actual observed short-term revenue would be biased on the low end compared to what a revenue maximizing owner would be likely able to accomplish.

@@

2. Potential ($REV_p$)}: Observed bookings extrapolated to a full year plus imputed bookings during observed blocked days.  Our attempt here is to represent the potential revenue if the property had been listed on Airbnb all year long.  The imputed bookings are imputed by comparing the property specific performance during the non-blocked days to overall market performance on those days and then applying that relative ratio to the overall market performance on the extrapolated and blocked days.

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Potential' revenue metric to be a better estimate of market performance. We chart the distribution of the Actual ($REV_a$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 6). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Actual short-term revenues range from neglible to highs of greater than \$60,000, with a collection of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is shifted slightly to the right due to our time extrapolation and blocked days imputation.

While this simple analysis does not control for location and product differences, it does highlight the fact that some profit seeking Airbnb owners are likely making greater revenues than in long-term rentals, but that many are not. The high number of very low yield Airbnb properties suggests that some owners may not be focused on maximizing revenue, but instead prefer the flexibility of short-term leasing.^[Do note that we have not considered any overall market demand issues when analyzing potential revenue.]

@@

To limit the potential impacts of systematic bias from omitted variables or misspecification in our hedonic model on later analyses, we've also used a matched-pairs or clustering approach as the other half of our ensemble.  In this method, we first group properties based on structure type, bedroom and bathroom combination and then find the geographically nearest 7 long-term observations. For example, all 2-bed, 2-bath apartments are grouped.  For each short-term rental in this group we find the nearest 7 long-term rental comparables (geographically), many of which are in the same building.  We compute the median of these seven comparables as our matched pairs estimation.

We then average the hedonic model and matched pairs prediction for each short term property to create an estimate of its long-term rental rate. As a validation check, we find some variation between the predictions from the two models for specific properties, however, the overall correlation between the two estimates is 97.6%.

@@

Overall, 57.3%^[If we just look at actual revenue, ignoring blocked days and properties with less than a full 12 months of listing, the figure is 42.5%] of the 2,415 short-term urban rental properties in our sample were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

We specify three, increasingly complex, logistic regression models.

@@

For Model 2 we add the four geographic submarkets, $SubMrkt$ to better control for geographic variation in the probabilities.

@@

where $GPR$ is the total number of guests allowed per bedroom^[Hosts set the total number of guests.  We have normalized this by bedroom to avoid co-linearity in our model with the bedroom variables.], $MinStay$ is the minimum number of nights that a stay may have and $CancelPolicy$ is the difficulty with which a booking may be cancelled.  We have grouped the cancellation policies into Flexible (24 hour notice), and Strict (those requiring five or more days notice, often with less than 100% refunds).

@@

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units (reference category) show greater likelihood of short-term preference than 2 bed / 2 bath combinations. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit. In other words, larger units are equally desirable to tourists and local residents alike.

@@

Finally, in Model 3 we add in host policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Finally, flexible cancellation policies result in a considerably lower probability of short-term preference. Strict cancellation may be a luxury available only for the best and most highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sensitivity tests. The most influential assumptions that we make are in regards to; 1) unobserved costs of short term leasing (furnishings, additional maintenance, turnover, etc.) and; 2) that the properties in the two markets in our datasets are equivalent in terms of unobserved variables, such as quality and condition. We chose these sensitivity tests so that we consider situations that each favor one of leasing strategy options. To approximate how these assumption may be influencing our results, we re-estimate our models under two new conditions:

@@

Table 3 shows the results of our sensitivity testing, with the original model results shown for comparison.  For the five variables with the strongest relationship in the original model -- House, City, Suburban, Guests per Bedroom and Flexible Cancel -- the signs of the coefficient remain identical across the sensitivity tests and the variation in magnitudes are small, relatively speaking.   A few of the originally less significant variables do show relatively larger deviations in coefficient magnitude, but in general, the models appear to be stable across the sentivity tests.  Do note that the intercept terms does change across models, which would be expected as this represents the base probability of short-term preference which we'd expect to move in concert with our changing assumptions.

@@

Comparing annual revenues from the two leasing strategies shows that revenues from more traditional long-term rentals are, on average, similar to those from Profit Seeking short term rentals, however, the long-term revenues are much less variable (Figure 6).  The long right tail to the distribution of short-term revenues suggesting that some owners are doing better, and occasionally much better, in the short-term market than they would on the long-term market. This univariate comparison of revenues does not tell the entire story, as the Airbnb market does not have the same spatial distribution as the long-term market (Figure 4). Simply put, short-term leasing is much more common in the central areas of the city, as would be expected.

In comparing the short-term observations with estimated long-term rents we found that just over half of our selected Airbnb properties brought in more revenue with its short-term leasing strategy than it would have with a long-term approach. As expected, this probability is not equal across space or property type.  Apartments are much more likely than houses to be more profitable as a short-term rental. City-Core properties are more likely than other properties to prefer a short-term strategy. For instance, 65\% of apartments in the city core generated more revenue on Airbnb than they likely would have on the long-term market while only 23\% of suburban houses did.

The logistic regression models (Tables 2 and 3) highlight which characteristics lead to increased short term preference.  Small apartments in the central city are the best bet for generating greater revenues through short-term leasing.  More flexibility around the number of guests and the minimum stay increase the probability of short-term preference, while the reverse is true of cancellation policies (more strict equals higher short-term preference probability).  As a host, it appears to be best to do what you can to attract bookings and then make it difficult for those bookings to be cancelled.

Testing for the sensitivity of the initial results through changes two of our core assumptions -- short term leasing costs and sample similarity between LTR and STR properties -- we find the qualitative finding to be generally robust.  While the base probability of short-term preference changes with changing assumptions (as expected), the sign and magnitudes of the co-variates remains relatively robust to changes in key assumptions.
